target_offtarget

### 2025-01-21 manuscript_figures.Rmd
updating figures

## >>>> Figure 1: plots of TE experiment data - read counts and read depths

@@

read_count <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(Viral Reads)")

@@

read_count_norm1 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads)")

@@

read_count_norm2 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)")

@@

read_count_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(Viral Reads)") +
  xlab("Viral load")

@@

read_count_norm2_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)") +
  xlab("Viral load")

### ***Final plot***
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.height=6, fig.height=8}
#plot just the two lefthand panels from the original figure

@@

ggsave("figures/manuscript_figure_2025/PDF/Figure_1.pdf",
       width=6,
       height=8)

ggsave("figures/manuscript_figure_2025/PNG/Figure_1.png",
       width=6,
       height=8)

@@

read_depths <- 
  depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(mean_depth),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(mean read depth)")

@@

### ***Final plot***
```{r}

@@

@@

### Reaed-in data / Make plots
```{r}
#metadata

@@

pool_readcounts <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = virus, y = log(matched))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (Viral Reads)") + xlab("Viral load (copies / ml)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_counts_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (Viral Reads)") + xlab("Viral load (copies / ml)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_readcounts_norm <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = virus, y = log(norm_counts1))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log(viral reads/cleaned reads)") +  xlab("Viral load (copies / ml)")

pool_norm1_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log(viral reads/cleaned reads)") +   xlab("Viral load (copies / ml)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_readcounts_norm2 <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = virus, y = log(norm_counts2))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)") + xlab("Viral load (copies / ml)")

pool_norm2_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
  xlab("Viral load (copies / ml)")

@@

pool_readdepths <- 
  depths_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = virus, y = log(mean_depth))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (mean read depth)") + xlab("Viral load (copies / ml)")

pool_depths_sum <- 
  depths_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(mean_depth),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    # axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (mean read depth)") +
  xlab("Viral load (copies / ml)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

@@

#ggsave("figures/manuscript_figures_pdf/FigureS2.pdf",width=10,height=7)

@@

total_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(total_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
  ggtitle("Total Reads") +
  ylab("Log(Total Reads)") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

viral_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(viral_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
  ggtitle("Total Viral Reads") +
  ylab("Log(Total Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

ATCC_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(ATCC_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log()) + xlab("Viral load (copies / ml)") +
  ggtitle("Spike In Viral Reads") +
  ylab("Log(Spike In Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

>>>>

### Figure 1 - linear model

depths_reads_sub <- depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  mutate(log_depth = log10(mean_depth)) |>
  mutate(log_viral_load = log10(Viral.load))

virus <- (unique(depths_reads_sub$virus)) %>%
  rep(., each = 2) %>%
  data.frame()

lm_tidy <- 
  map(list_models, broom::tidy) %>%
  do.call(rbind.data.frame, .) %>%
  cbind(virus, .) %>%
  rename(virus = ".") %>%
  select(virus, term, estimate) %>%
  pivot_wider(names_from = "term", values_from = "estimate")

#### Make plot
```{r}
ggplot() +
  geom_point(data = depths_reads_sub,
             mapping = aes(log_viral_load, log_depth, color = virus)) +
  geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load), linetype="dotted") +
  geom_text(data = lm_combined,
            colour = "black",
            aes(
              x = 4.7,
              y = 3,
              label = round(Coeff_x1, digits = 2)),
            size = 3) +
  geom_text(data = lm_combined,
            colour = "black",
            aes(x = 4, 
                y = 3, 
                label = "Slope ="),
            size = 3) +
  geom_text(data = lm_combined,
            colour = "black",
            aes(
              x = 4.7,
              y = 2,
              label = round(R2, digits = 2)),
            size = 3) +
  geom_text(data = lm_combined,
            colour = "black",
            aes(x = 4, 
                y = 2, 
                label = "R2 ="),
            size = 3) +
  facet_wrap( ~ virus, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray"),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
  xlab("log10(Viral Load)") +
  ylab("log10(Mean Read Depth)") +
  scale_color_manual(values = cols) +
  guides(color = FALSE)

#### Final plot
```{r}
#ggsave("figures/compare_spike_ins_atcc/model_readdepth_dedup.png")

ggsave("figures/manuscript_figure_2025/PDF/Figure_1.pdf",
       width = 7.5,
       height = 4)

ggsave("figures/manuscript_figure_2025/PNG/Figure_1.png",
       width = 7.5,
       height = 4)

## >>>> now redundant? - Figure 1: plots of TE experiment data - read counts and read depths

@@

read_count <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("log10(Viral Reads)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

@@

read_count_norm1 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("log10(viral reads/cleaned reads)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

@@

read_count_norm2 <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = (Viral.load),
    y = log10(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log (viral reads/cleaned reads/genome length)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

@@

read_count_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE, method = "lm") +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log (Viral Reads)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

@@

read_count_norm2_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE, method = "lm") +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log (viral reads/cleaned reads/genome length)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
```

### ***Final plot***
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=10}
#plot just the two lefthand panels from the original figure

@@

ggsave("figures/manuscript_figure_2025/PDF/Figure_S1.pdf",
       width=7,
       height=8)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S1.png",
       width=7,
       height=8)

@@

read_depths <- 
  depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(mean_depth),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log (mean read depth)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

@@

### ***Final plot***
```{r, warning = FALSE, echo = FALSE, message = FALSE}

@@

ggsave("figures/manuscript_figure_2025/PDF/Figure_S1.pdf",
       width = 12,
       height = 8)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S1.png",
       width = 12,
       height = 8)

@@

### Read-in data / Make plots
```{r}
#metadata

@@

pool_counts_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log10(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (Viral Reads)") + xlab("Log (Viral load)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
  scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5})))

# pool_readcounts_norm <- 
#   counts_pool |>
#   filter(Background != "p6") |>
#   filter(Background != "control") |>
#   ggplot(aes(x = virus, y = log10(norm_counts1))) +
#   geom_boxplot() +
#   facet_grid(Pool ~ type) +
#   theme_few() +
#   theme(
#     axis.title.x = element_blank(),
#     # axis.text.x = element_blank(),
#     axis.title.y = element_text(size = 10),
#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
#   ) +
#   ylab("log10(viral reads/cleaned reads)")

pool_norm1_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log10(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray") 
  ) +
  ylab("log10(viral reads/cleaned reads)") + xlab("Log (Viral load)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
    scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5})))

# pool_readcounts_norm2 <- 
#   counts_pool |>
#   filter(Background != "p6") |>
#   filter(Background != "control") |>
#   ggplot(aes(x = virus, y = log10(norm_counts2))) +
#   geom_boxplot() +
#   facet_grid(Pool ~ type) +
#   theme_few() +
#   theme(
#     axis.title.x = element_blank(),
#     # axis.text.x = element_blank(),
#     axis.title.y = element_text(size = 10),
#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
#   ) +
#   ylab("log10(viral reads/cleaned reads/genome length)")

pool_norm2_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log10(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("log10(viral reads/cleaned reads/genome length)") + xlab("Log (Viral load)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
    scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5})))

@@

# pool_readdepths <- 
#   depths_pool |>
#   filter(Background != "p6") |>
#   filter(Background != "control") |>
#   ggplot(aes(x = virus, y = log10(mean_depth))) +
#   geom_boxplot() +
#   facet_grid(Pool ~ type) +
#   theme_few() +
#   theme(
#     axis.title.x = element_blank(),
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     axis.title.y = element_text(size = 10),
#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
#   ) +
#   ylab("Log (mean read depth)")

pool_depths_sum <-
  depths_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log10(mean_depth),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (mean read depth)") +
  xlab("Log (Viral load)") + 
  scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5}))) +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

@@

ggsave("figures/manuscript_figure_2025/PDF/Figure_S2.pdf",
       width = 10,
       height = 7)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S2.png",
       width = 10,
       height = 7)

@@

total_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(total_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
  ggtitle("Total Reads") +
  ylab("log10(Total Reads)") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

viral_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(viral_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
  ggtitle("Total Viral Reads") +
  ylab("log10(Total Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

ATCC_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(ATCC_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10(labels = scales::label_log10()) + xlab("Viral load (copies / ml)") +
  ggtitle("Spike In Viral Reads") +
  ylab("log10(Spike In Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))
